Algebraic File Synchronization: Adequacy and Completeness

### 2017-10-08 p2/axioms.tex
Remove unused symbols; clarify text
Improvements

% TODO \nequiv\cbrk -> \wrks

@@

\begin{mydef}[$A\indep B$: Independent commands, sequences and sets of commands]\label{def_indep}
Two commands $\cxynv$ and $\czwmv$ 
are independent, and we write $\cxynv\indep\czwmv$ if 
they commute and do not break all filesystems:
\[ \cxynv\cc\czwmv \equiv \czwnv\cc\cxynv \nequiv \cbrk. \]
For two sequences or unordered sets of commands $A$ and $B$ we write $A\indep B$ if
for all $\alpha$ in $A$ and all $\beta$ in $B$, $\alpha\indep\beta$.
We also write $\alpha\indep B$ iff $\{\alpha\}\indep B$.
\end{mydef}

@@

We proceed in two steps. First, we note that
from \cref{ax_separate_commute,ax_separate_nobreaks} we know that
if $n\unrel m$, then $\cxynv$ and $\czwmv$ must necessarily be independent.

We prove the second step by contradiction.
We assume that both $\cxynv\indep\czwmv$ and $n\nunrel m$,
and show that it leads to contradiction.
As $n\nunrel m$
we know that $n=m$, $n\descendant m$ or $m\descendant n$.

>>>>

@@

\begin{mydef}[$A\indep B$: Independent commands, sequences and sets of commands]\label{def_indep}
Two commands $\cxynv$ and $\czwmv$ 
are independent, and we write $\cxynv\indep\czwmv$ if 
they commute and do not break all filesystems:
\[ \cxynv\cc\czwmv \equiv \wrksx{\czwnv\cc\cxynv}. \]
For two sequences or unordered sets of commands $A$ and $B$ we write $A\indep B$ if
for all $\alpha$ in $A$ and all $\beta$ in $B$, $\alpha\indep\beta$.
We also write $\alpha\indep B$ iff $\{\alpha\}\indep B$.
\end{mydef}

@@

For the right-to-left part, we note that
from \cref{ax_separate_commute,ax_separate_nobreaks} we know that
if $n\unrel m$, then $\cxynv$ and $\czwmv$ must necessarily be independent.

We prove the left-to-right part by contradiction.
We assume that both $\cxynv\indep\czwmv$ and $n\nunrel m$,
and show that it leads to contradiction.
As $n\nunrel m$
we know that $n=m$, $n\descendant m$ or $m\descendant n$.

### 2017-10-08 p2/def-commands.tex
Improvements

\noindent
For reasons also listed in \cite{NREC}, in this paper we will not consider
a $move$ or $rename$ command. Regarding the theoretical reasoning we aim to follow,
this turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for $move$ would call for a more complicated model.
From a pragmatic perspective, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by pre- and post-processing changes in the filesystem to
discover them, which can enhance usability, but which
(especially when a rename is combined with changes to the content)
is a non-trivial problem in itself.

>>>>

\noindent
For reasons also listed in \cite{NREC}, in this paper we will not consider
a $move$ or $rename$ command.
This turns out to be useful because this would be the only command that affects
filesystems at two nodes at once, therefore describing 
the dependencies for $move$ would call for a more complicated model.
From a pragmatic perspective, this restriction does not mean that in an application
implementing conflict resolution using the algorithm described here would not be
able to handle renames by pre- and post-processing changes in the filesystem to
discover them, which can enhance usability, but which
(especially when a rename is combined with changes to the content)
is a non-trivial problem in itself.

### 2017-10-08 p2/def-filesystem.tex
F depends on N
Improved text. Added definitions of construction/destruction commands.
Improvements

We model a working filesystem
using a function $\FS$ with a set of nodes (filesystem paths) $\setn$ as its domain,
and a set of possible contents or values $\setv$ as its codomain.
\begin{mydef}[Filesystems, $\setn$, $\setv$, $\setfs$ and $\empt$]
\[ \FS: \setn \rightarrow \setv \]
In our model, $\setn$ 
serves as a namespace or ``skeleton'' for the filesystem.
It contains all possible nodes, including the ones 
where the file system contains no file or directory,
where in our model $\FS$ has a special value, $\empt\in\setv$.
We write $\setfs$ for the set of filesystems.
\end{mydef}

@@

As usual, we write $[v]$ for the equivalence class of $v$, which represents its type.
We also consider metadata to be part of the values in $\setv$.

The formal definition of the tree property is therefore
\begin{mydef}[Tree property]
\[ \forall n\in\setn, \FS\in\setfs: \FS(n) \neq \empt \Longrightarrow \FS(\parent(n)) \in \setd \]
where $\parent(n)$ is defined, and where $\setd\subset\setv$ is the set of directories.
\end{mydef}

@@

% This means that as we move down from the root of a tree of nodes,
% the types of values we encounter in the filesystem can only change according to the
% transition diagram in \cref{fig_transition}, from directories ($\cchard$) to files ($\ccharf$)
% to empty nodes ($\ccharb$).
% 
% \begin{figure}[htb]
% \input{fig-type-transition}
% \caption{Transitions between types of values from the root of a tree in a filesystem}\label{fig_transition}
% \end{figure}

In this paper, $\FS$ and $\GS$ denote filesystems,
and $n$, $m$ and $o$ are nodes in $\setn$.
We write $\valf$ for an arbitrary file value. % element in $\setf$.

>>>>

We model a working filesystem
using a function $\FS$ with a set of nodes (filesystem paths) $\setn$ as its domain,
and a set of possible contents or values $\setv$ as its codomain.
We consider metadata to be part of the values in $\setv$.
\begin{mydef}[Filesystems, $\setn$, $\setv$, $\setfs$ and $\empt$]
\[ \FS: \setn \rightarrow \setv \]
In our model, $\setn$ 
serves as a namespace or ``skeleton'' for the filesystem.
It contains all possible nodes, including the ones 
where the file system contains no file or directory,
where in our model $\FS$ has a special value, $\empt\in\setv$.
We write $\setfs$ for the set of filesystems over $\setn$.
\end{mydef}

@@

As usual, we write $[v]$ for the equivalence class of $v$, which represents its type.

The formal definition of the tree property is therefore
\begin{mydef}[Tree property]
\[ \forall n\in\setn, \FS\in\setfs: \FS(n) \neq \empt \Longrightarrow \FS(\parent(n)) \typeeq \vald \]
where $\parent(n)$ is defined, and where $\vald\in\setv$ is a directory value.
\end{mydef}

@@

In this paper, $\FS$ and $\GS$ denote filesystems,
and $n$, $m$ and $o$ are nodes in $\setn$.
We write $\valf$ for a file value. % element in $\setf$.

### 2017-10-08 p2/ordering.tex
Improvements

\begin{mycor}\label{subseq_in_orderset}
For simple sequences $S$ and $T$,
if $T\in\orderset{S}$, then for any subsequence of $S$, $S_0$, which contains
only part of the commands in $S$ but in the same order as in $S$,
and for the corresponding subsequence of $T$, $T_0$, which
contains the same commands but in the order they are found in $T$,
$T_0\in\orderset{S_0}$ must hold.
\end{mycor}
\begin{proof}
This is because
if there are no command pairs in $T$ not conforming to \cref{def_orderset},
then there cannot be any in $T_0$.
\end{proof}

\begin{mylem}\label{simple_distant_pairs}
For any simple sequence $\wrksx{S}$,
if $\cxynv\cc\czwmv$ is a subsequence of $S$ and $n=\parent(m)$ or $\parent(n)=m$,
then $\wrks{\cxynv\cc\czwmv}$.
\end{mylem}
\begin{proof}
Informally speaking, the proposition means that
the two commands must form a construction or destruction pair even
if they are not next to each other.
This is true because in simple sequences
there cannot be another command on $n$ or $m$,
so if the commands are incompatible, no command between them can change that.
In fact, this \namecref{simple_distant_pairs} is the general case
of \cref{ax_directchild_breaks,ax_directparent_breaks}.

@@

\begin{mycor}\label{order_is_only_possible}
For any simple sequence $S$ and its permutation $S'$, 
\[ S'\not\in\orderset{S} \Longrightarrow S'\equiv\cbrk. \]
\end{mycor}
\begin{proof}
This follows directly from \cref{simple_distant_pairs}.
If $S'\not\in\orderset{S}$ then by \cref{def_orderset}
it contains two commands on a node $n$ and its parent
that placed next to each other do not form
a construction or destruction pair.
Let them be $\cxynv$ and $\czwmv$.
By \cref{ax_directchild_breaks,ax_directparent_breaks}
we know
$\cxynv\cc\czwmv\equiv\cbrk$, which is a subsequence of $S'$,
and therefore by \cref{simple_distant_pairs} $S'\equiv\cbrk$.
\end{proof}

@@

>>>>

\begin{mycor}\label{subseq_in_orderset}
For simple sequences $S$ and $T$,
if $T\in\orderset{S}$, then for any subsequence of $S$, $S_0$,
and for the corresponding subsequence of $T$, $T_0$, which
contains the same commands but in the order they are found in $T$,
$T_0\in\orderset{S_0}$ must hold.
\end{mycor}
\begin{proof}
This is because
if there are no command pairs in $T$ not conforming to \cref{def_orderset},
then there cannot be any in $T_0$.
\end{proof}

\begin{mylem}\label{simple_distant_pairs}
For any simple sequence $\wrksx{S}$,
if $\cxynv\cc\czwmv$ is a subsequence of $S$ and $n=\parent(m)$ or $\parent(n)=m$,
then $\wrksx{\cxynv\cc\czwmv}$.
\end{mylem}
\begin{proof}
Informally speaking, the proposition means that
the two commands must form a construction or destruction pair even
if they are not next to each other.
This is true because in simple sequences
there cannot be another command on $n$ or $m$,
so if the commands are incompatible, no command between them can change that.
In fact, this \namecref{simple_distant_pairs} is the general case
of \cref{ax_directchild_breaks,ax_directparent_breaks}.

@@

\begin{mycor}\label{order_is_only_possible}
For any simple sequence $S$ and its permutation $S'$, 
\[ S'\not\in\orderset{S} \Longrightarrow S'\equiv\cbrk. \]
\end{mycor}
\begin{proof}
This follows directly from \cref{simple_distant_pairs}.
If $S'\not\in\orderset{S}$ then by \cref{def_orderset}
it contains two commands on a node $n$ and its parent
that placed next to each other do not form
a construction or destruction pair.
Let them be $\cxynv$ and $\czwmv$.
By \cref{ax_directchild_breaks,ax_directparent_breaks}
we know
$\cxynv\cc\czwmv\equiv\cbrk$, which is a subsequence of $S'$,
and therefore by \cref{simple_distant_pairs}, $S'\equiv\cbrk$.
\end{proof}

@@

\begin{mydev}[$\orderd{S}$]
It follows from \cref{simple_reorder_equiv} that given a simple set of commands $S$,
which can be applied to a filesystem without breaking it,
all sequences in $\orderset{S}$
represent the same partial endofunction on filesystems.
This allows us to treat such a set as a sequence for any purpose
where the internal order of the commands is irrelevant.
We write $\orderd{S}$ to represent this endofunction.
\end{mydev}

### 2017-10-08 p2/works-inverse.tex
Improvements

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.
We will write $\Dom{\sqs{A}}$ for $\bigcap_{A\in\sqs{A}} \Dom{A}$,
and $\sqs{A}\cc\sqs{B}$ for $\{A\cc B\whr A\in\sqs{A}, B\in\sqs{B}\}$.
We will also use $A\cc\sqs{B}$ with the meaning $\{A\}\cc\sqs{B}$.

@@

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\worksmnil{\{A,B\}}$,
and there are commands on node $n$ in both $A$ ($\cxynv\in A$) and $B$ ($\czwnv\in B$),
then the input types of these commands must match ($[x]=[z]$).
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
is easily shown using a proof by contradiction: if $[x]\neq [z]$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}

@@

By definiton, for all objects where it is defined, $a\T\typeeq a$.
As representatives of filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts, 
all propositions proven so far apply to them as well.

\begin{mycor}\label{cor_typemap_transitive}
For any filesystem $\FS$ and command $\alpha$, 
$\alpha\T(\FS\T) = (\alpha\aFS)\T$.
TODO needed?
\end{mycor}

@@

% TODO \alpha \neq \cbrk
\begin{mylem}\label{repr_comm_inject}
Representative commands $\alpha\T$ are partial bijections over representative filesystems ($\FS\T$).
Because partial functions are trivially surjective when restricted to their images, this is
equivalent to saying that representative functions are injective except for cases
where they are not defined:
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
Choose values that satisfy the left side, and let $\alpha\T = \cTxynv$.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\FS\T$. 
The only remaining case therefore is where they have the same values
at all nodes except at $n$.
As $\cTxynv(\FS\T)\neq\fsbroken$, we know $[\FS\T(n)]=[x]$.
We also know that $\GS\T(n)\neq\FS\T(n)$, and therefore, as there is a one-to-one correspondence between values and types,
$[\GS\T(n)]\neq[x]$.
This implies $\cTxynv(\GS\T)=\fsbroken$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

@@

We write ${S\T}^{-1}$ for the inverse of the sequence $S\T$, which consists of the inverses of the commands in $S\T$
in reverse order.
\end{mydef}
Per definition, for any filesystem $\FS\T$ outside the image of $\alpha\T$, the inverse $\alpha\T^{-1}$ is not defined.

>>>>

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.

We will write $\Dom{\sqs{A}}$ for $\bigcap_{A\in\sqs{A}} \Dom{A}$,
and $\sqs{A}\cc\sqs{B}$ for $\{A\cc B\whr A\in\sqs{A}, B\in\sqs{B}\}$.
We will also use $A\cc\sqs{B}$ with the meaning $\{A\}\cc\sqs{B}$.

@@

\begin{mylem}\label{worksinputmatch}
If $S$ and $T$ are minimal sequences, $\worksmnil{\{A,B\}}$,
and there are commands on node $n$ in both $S$ ($\cxynv\in A$) and $T$ ($\czwnv\in B$),
then the input types of these commands must match ($[x]=[z]$).
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
is easily shown using a proof by contradiction: if $[x]\neq [z]$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $S$ and $T$ cannot work on the same filesystem.
\end{proof}

@@

By definiton,
for any filesystem $\FS$ and command $\alpha$, 
$\alpha\T(\FS\T) = (\alpha\aFS)\T$.

% TODO ???
Also, for all objects where it is defined, $a\T\typeeq a$.
As representatives of filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts, 
all propositions proven so far apply to them as well.

@@

% TODO \alpha \neq \cbrk
\begin{mylem}\label{repr_comm_inject}
Representative commands $\alpha\T$ are partial bijections over representative filesystems
($\FS\T$, $\GS\T$).
Because partial functions are trivially surjective when restricted to their images, this is
equivalent to saying that representative functions are injective except for cases
where they are not defined:
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
Choose values that satisfy the left side, and let $\alpha\T = \cTxynv$.
If $\FS\T$ and $\GS\T$ have different values (types) at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\FS\T$. 
The only remaining case therefore is where they have the same values
at all nodes except at $n$.
As $\cTxynv(\FS\T)\neq\fsbroken$, we know $[\FS(n)]=[x]$.
We also know that $\GS\T(n)\neq\FS\T(n)$, and therefore
$[\GS(n)]\neq[x]$.
This implies $\cTxynv(\GS\T)=\fsbroken$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

@@

We write ${S\T}^{-1}$ for the inverse of the sequence $S\T$, which consists of the inverses of the commands in $S\T$
in reverse order.
\end{mydef}
By definition, for any representative filesystem $\FS\T$ outside the image of $\alpha\T$, the inverse $\alpha\T^{-1}$ is not defined.
